% !TeX spellcheck = en_GB
%!TEX root = ../thesis.tex


\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

This work is a result of many contributions from people who have helped me along the way and who I am very grateful to.

First, I want to thank my supervisor, Oded Zilberberg. He gave me every opportunity to pursue my own ideas and the guidance necessary for their fruition. His mentorship was essential for me to navigate the world of academic physics and get the most out of my time as a doctoral student. I am also indebted to Chitra Ramasubramanian and Alex Eichler, who provided a lot of support in the early stages of this work when it was needed most. 

During my stay at the ITP, I had the fortune of getting to know many wonderful people. I am especially grateful the QUEST group for providing a relaxed, friendly environment. Their office camaraderie was an indispensable ingredient of my PhD experience and will be sorely missed. In particular, I would like to thank my closest collaborators: Toni Heugel, who was a patient guide through the realm of nonlinearity, Anina Leuch, with whom I set out to question undergraduate textbooks, and Javier del Pino, who kickstarted the project that ended up being a turning point in my professional life.

I cannot thank my girlfriend Tereza enough. She was with me through every high and low of this work, had faith in me when I had none and always reminded me of what really matters in life. Finally, my deepest gratitude goes to my mother. Without her unconditional support, none of this would ever have been possible.